---
title: "Final_Code_OAI"
author: "Reshani Abayasekara"
date: "2024-06-25"
output: html_document
---
```{r Packages Used}
module load gcc/8.3.0 r/4.1.2
R
library(mmrm)
library(pander)
library(tidyverse)
library(registr)
library(reportRmd)
library(cowplot)
library(factoextra)
library(Hmisc)
library(gridExtra)
library(cluster)
library(fpc)
library(pROC)
library(datawizard)
library(pracma)
library(GLMMadaptive)
setwd(Sys.getenv("SCRATCH"))
setwd(("/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac"))
setwd("/gpfs/fs0/scratch/o/oespinga/reshani0/")
```

Accelerometer data available
```{r OAI data}
min8=read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/Acceldatabymin08.txt", sep="|", header=T) 
min6=read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/Acceldatabymin06.txt", sep="|", header=T) 
#accel6=read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/Accelerometry06.txt", sep="|", header=T) 
#accel8=read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/Accelerometry08.txt", sep="|", header=T)
```

```{r Files geneerated as part of cleaning}
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData") #This is the outcomes data but it must be called as OAI_KL_JSW_data that's how to work with it.
load("outcomes_v68.RData")#This is the outcomes data for visit 6 and visit 8
load("unreg_times_v8.Rdata") #Unregistered data
load("unreg_times_v6.Rdata")#unregistered data
load("combo_test_v6.Rdata")#AURC data with outcomes for visit 6 
load("combo_test_v8.Rdata")#AURC data with outcomes for visit 8
load('visits_ID.Rdata')#ID list of sample pop
```

Sample population determination 
```{r Sample population}
#I am interested in those who have 4-7 valid days:
#Finding those who qualified to have accelerometer data collected but chose not to at each visit. 
notparticipating_accel6=filter(accel6,V06APASTAT=="Not participating") #585 are eligible for the study but chose not to participate
notparticipating_accel8=filter(accel8,V08APASTAT=="Not participating")#276 are eligible for the study but chose not to participate 

#Finding those who participated at each visit with 4 or more valid days:
accel6=filter((filter(accel6,V06APASTAT=="Worn with data")),V06ANVDAYS>=4)
accel8=filter((filter(accel8,V08APASTAT=="Worn with data")),V08ANVDAYS>=4)
save(accel6, file = 'accel6.Rdata') #These are the individuals present at v6 with more than 4 valid days of accelerometer data
save(accel8, file = 'accel8.Rdata') #These are the individuals present at v8 with more than 4 valid days of accelerometer data

#How many ppl are present at both visits: 
combined_visit=inner_join(x=accel6,y=accel8, by='ID') #1343 people present at both
visits_ID=data.frame(ID=combined_visit$ID) #IDs of people present at both 
save(visits_ID, file = 'visits_ID.Rdata') ##This is my ID's of the sample pop
```

Initial data exploration:
```{r Number of obese individuals}
combo<-inner_join(OAI_KL_JSW_data,visits_ID,by="ID")

combo_edit<-combo%>%
  dplyr::filter(VISIT=="V00")
#I see that we have 465 people who are obese, 540 who are overweight, 335 who are healthy and 3 who are underweight

combo_edit<-combo%>%
  dplyr::filter(VISIT=="V06")%>%
  transform(obese_over=as.factor(ifelse(BMI>=25,"yes","no")))
#We see that 470 are obese and 871 are not obese while 2 did not have valid BMI readings.
#We see that 987 are obese or overweight while 354 are not while 2 did not have valid BMI readings.
```

```{r Missing Values}
#From the sample population how many ppl are missing activity count values?
#Visit6:
min6_pop<-min6%>%
  dplyr::filter(ID %in% visits_ID$ID)
sum(is.na(min6_pop$V06MINCnt))
min6_pop=min6_pop %>%  
  group_by(ID) %>%  
  summarise(Missing=sum(is.na(V06MINCnt)), .groups="drop")%>%
  dplyr::filter(Missing>0)
##We see that ID 9718951 had 15 missing counts back to back and ID 9732751 had 1 missing count.
#Visit8:
min8_pop<-min8%>%
  dplyr::filter(ID %in% visits_ID$ID)#gets us n=1343
sum(is.na(min8_pop$V08MINCnt))#73 missing values across the 4-7 days
min8_pop=min8_pop %>%  
  group_by(ID) %>%  
  summarise(Missing=sum(is.na(V08MINCnt)), .groups="drop")%>%
  dplyr::filter(Missing>0)
#We see that ID 9401202  had 73 missing counts over their person Notice that they are back to back 

##Let's replace the missing values with 0
min6=min6%>%
  dplyr::filter(ID %in% visits_ID$ID)
min6[is.na(min6)] <- 0
save(min6,file="min6.Rdata")#Min-min accel data for our sample pop that is cleaned 

min8=min8%>%
  dplyr::filter(ID %in% visits_ID$ID)
min8[is.na(min8)] <- 0
save(min8,file="min8.Rdata")#Min-min accel data for our sample pop that is cleaned 
```
Median vs mean day
```{r Bland Altman plot}
#Using the cleaned data
load("min6.Rdata")
load("min8.Rdata")

#Accelerometer data at visit 6 for all subjects present at both visit 6,8 
#Mean and median day:
min6_day<-na.omit(min6)%>%
  group_by(ID,V06MinSequence)%>%
  summarise(avg_count_mean=mean(V06MINCnt),
            avg_count_median=median(V06MINCnt))

#create new column for average measurement
min6_day$avg <- rowMeans(min6_day[,c("avg_count_mean", "avg_count_median")], na.rm=TRUE)

#create new column for difference in measurements
min6_day$diff <- min6_day$avg_count_mean - min6_day$avg_count_median

#view first six rows of data
head(min6_day)

#find average difference
mean_diff <- mean(min6_day$diff)

#find lower 95% confidence interval limits
lower <- mean_diff - 1.96*sd(min6_day$diff)
upper <- mean_diff + 1.96*sd(min6_day$diff)

#load ggplot2
library(ggplot2)
#create Bland-Altman plot
bplot_v6<-ggplot(min6_day, aes(x = avg, y = diff)) +
  geom_point(size=2) +
  geom_hline(yintercept = mean_diff) +
  geom_hline(yintercept = lower, color = "red", linetype="dashed") +
  geom_hline(yintercept = upper, color = "red", linetype="dashed") +
  labs(title="48-month",x = "Average activity count", y = "Difference mean and median activity count",tag = "A")+
  theme(text = element_text(size = 20))+
  theme(axis.title=element_text(size=20))+
  theme(legend.text=element_text(size=18),legend.title=element_text(size=20))+
  theme(plot.title = element_text(size=22,face="bold"))

##Visit 8 Bland Altman Plot

min8_day<-na.omit(min8)%>%
  group_by(ID,V08MinSequence)%>%
  summarise(avg_count_mean=mean(V08MINCnt),
            avg_count_median=median(V08MINCnt))

#create new column for average measurement
min8_day$avg <- rowMeans(min8_day[,c("avg_count_mean", "avg_count_median")], na.rm=TRUE)

#create new column for difference in measurements
min8_day$diff <- min8_day$avg_count_mean - min8_day$avg_count_median

#view first six rows of data
head(min8_day)

#find average difference
mean_diff <- mean(min8_day$diff)

#find lower 95% confidence interval limits
lower <- mean_diff - 1.96*sd(min8_day$diff)
upper <- mean_diff + 1.96*sd(min8_day$diff)

#load ggplot2
library(ggplot2)
#create Bland-Altman plot
barplot_v8<-ggplot(min8_day, aes(x = avg, y = diff)) +
  geom_point(size=2) +
  geom_hline(yintercept = mean_diff) +
  geom_hline(yintercept = lower, color = "red", linetype="dashed") +
  geom_hline(yintercept = upper, color = "red", linetype="dashed") +
  labs(title="72-month",x = "Average activity count", y = "Difference mean and median activity count",tag = "B")+
  theme(text = element_text(size = 20))+
  theme(axis.title=element_text(size=20))+
  theme(legend.text=element_text(size=18),legend.title=element_text(size=20))+
  theme(plot.title = element_text(size=22,face="bold"))

grid.arrange(bplot_v6,barplot_v8,ncol = 1, nrow = 2)

```
After examining the bland altman plots we decided to move ahead with a median day. Below is how the median day was generated.
```{r Median Day}
load("min6.Rdata")#Cleaned min data
load("min8.Rdata")#Cleaned min data
min6_medday<-min6%>%
  group_by(ID,V06MinSequence)%>%
  summarise(value=median(V06MINCnt))%>%
  rename(id=ID)%>%
  rename(index=V06MinSequence)%>%
  dplyr::filter(id %in%visits_ID$ID)%>%
  as.data.frame()##This must be done to run registration
min6_medday <- replace(min6_medday, is.na(min6_medday), 0)

min8_medday<-min8%>%
  group_by(ID,V08MinSequence)%>%
  summarise(value=median(V08MINCnt))%>%
  rename(id=ID)%>%
  rename(index=V08MinSequence)%>%
  dplyr::filter(id %in% visits_ID$ID)%>%
  as.data.frame()##This must be done to run registration
min8_medday <- replace(min8_medday, is.na(min8_medday), 0)

```
Running curve registration:
```{r Curve registration}
#These are the saved files
min6_medday_reg<-register_fpca(Y=min6_medday,family = "gaussian",
                  npc = 2, 
                   Kt=8,
                    Kh=4,
                   gradient = FALSE,
                   warping = "piecewise_linear2")

save(min6_medday_reg,file="min6_medday_reg.Rdata")

min8_medday_reg<-register_fpca(Y=min8_medday,family = "gaussian",
                  npc = 2, 
                   Kt=8,
                    Kh=4,
                   gradient = FALSE,
                   warping = "piecewise_linear2")

save(min8_medday_reg,file="min8_medday_reg.Rdata")

```
Generating registration plots
```{r Legends}
#Plots
library(gghighlight)
library(grid)
load("min6_medday_reg.Rdata")
load("visits_ID.Rdata")
preds<-data.frame(pred_value=(min6_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min6_medday_reg$Y,preds)

##No control:
exc_control_ID<-OAI_KL_JSW_data%>%
  dplyr::filter(V00COHORT!="3: Non-exposed control group")%>%
  dplyr::filter(VISIT=="V00")%>%#we only need control IDs
  select(ID)%>%
  inner_join(visits_ID,by="ID")#sample pop in control group

#Replace ids with numbers
id_rep<-data.frame(id=exc_control_ID$ID,p_id=c(1:1337))
combo_rep<-inner_join(combo,id_rep,by="id")

test<-ggplot(combo_rep,aes(x=(tstar),y=value,col=p_id,group=p_id))+
  geom_line()+scale_color_gradient(low="blue", high="grey")+scale_color_viridis_c(na.value="white", limits = c(1, 1337),
                       breaks = c(1, 672,1337), 
                       guide = "colourbar")+ labs(x = "Unregistered time (Minute)", y = "Median activity count",
              tag = "A")+labs(color = "Participant Number")+
  scale_x_continuous(breaks=seq(0, 1440, 60))+ggtitle("48-month visit")+theme(axis.title=element_text(size=15))+theme(text = element_text(size = 15))+gghighlight(p_id==168,label_key=p_id,unhighlighted_params = list(colour = NULL, alpha = 0.3))+theme(plot.title = element_text(size=22,face="bold"))                                                            
```

```{r Curve Registration Plots}
library(gghighlight)
library(grid)
load("min6_medday_reg.Rdata")


preds<-data.frame(pred_value=(min6_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min6_medday_reg$Y,preds)

#Replace ids with numbers
id_rep<-data.frame(id=no_control_ID$ID,p_id=c(1:1343))
combo_rep<-inner_join(combo,id_rep,by="id")

pre_6reg<-ggplot(combo_rep,aes(x=(tstar),y=value,col=p_id,group=p_id))+geom_line(show.legend = FALSE)+labs(x = "Unregistered time (Minute)", y = "Median activity count",tag = "A")+labs(color = "Participant Id")+
  scale_x_continuous(breaks=seq(0, 1440, 120))+theme(text = element_text(size = 8))+theme(axis.title=element_text(size=8))+ scale_color_viridis_c()+ggtitle("48-month visit")+theme(plot.title = element_text(size=10,face="bold"))+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

#Warping
warping_6p_gaus<-ggplot(combo_rep, aes(x = tstar, y = (t_hat*1440), col = (p_id)))+
    geom_line(show.legend = FALSE) + labs(x = "Unregistered time (Minute)", y = "Registered time (Minute)",
              tag = "C")+labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_y_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 20))+theme(text = element_text(size = 10))+theme(axis.title=element_text(size=10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

##Post registration:
post_reg_gaus_6p<-ggplot(combo_rep,aes(x=(t_hat*1440),y=value,col=(p_id),group=p_id))+
    geom_line(show.legend = FALSE) +labs(x = "Registered time (Minute)", y = "Median Activity Count",tag = "E") +labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 10))+theme(text = element_text(size = 10))+theme(axis.title=element_text(size=10))+scale_color_viridis_c()+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

#Visit 8
library(gghighlight)
library(grid)
load("min8_medday_reg.Rdata")

preds<-data.frame(pred_value=(min8_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min8_medday_reg$Y,preds)

#Replace ids with numbers
id_rep<-data.frame(id=visits_ID$ID,p_id=c(1:1343))
combo_rep<-inner_join(combo,id_rep,by="id")

#Pre-registration curve
pre_8reg<-ggplot(combo_rep,aes(x=(tstar),y=value,col=p_id,group=p_id))+
    geom_line(show.legend = FALSE)+ labs(x = "Unregistered time (Minute)", y = "Median activity count",
              tag = "B")+labs(color = "Subject Id")+
  scale_color_gradient(low="blue", high="grey")+ scale_x_continuous(breaks=seq(0, 1440, 120))+theme(text = element_text(size = 20))+theme(text = element_text(size = 10))+theme(text = element_text(size = 10))+scale_color_viridis_c()+theme(axis.title=element_text(size=10))+ggtitle("72-month visit")+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+theme(plot.title = element_text(size=10,face="bold"))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

#Warping
warping_8p_gaus<-ggplot(combo_rep, aes(x = tstar, y = (t_hat*1440), col = (id)))+
    geom_line(show.legend = FALSE)+ labs( x = "Unregistered time (Minute)", y = "Registered time (Minute)",
              tag = "D")+labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_y_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

##Post registration:
post_reg_gaus_8p<-ggplot(combo_rep,aes(x=(t_hat*1440),y=value,group=p_id,col=(p_id)))+
    geom_line(show.legend = FALSE) +labs(x = "Registered time (Minute)", y = "Median Activity Count",
              tag = "F") +labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size =10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+ scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size =0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

mix_plot<-grid.arrange(pre_6reg,pre_8reg,
                  warping_6p_gaus, warping_8p_gaus,
                  post_reg_gaus_6p,post_reg_gaus_8p,
                  ncol = 2, nrow = 3)

ggsave(plot= mix_plot, filename = '/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/comboplot.pdf',width = 297,
  height = 210,
  units = c( "mm"))
```


Let's look at how many individuals met the recommended CDC guidelines at each visit.
```{r CDC guidelines v6}
mod_ac<-min6_medday_reg$Y%>% #The registered output for visit 6
  rename(ID=id)%>% 
  transform(t_hat=t_hat*1440)%>%
  dplyr::filter(between(value,2020,5998))%>%
  group_by(ID)%>%
  mutate(t_hatdiff=t_hat-lag(t_hat))%>%#Find the difference between the minute recorded before which had mod activity. This was done to see if the times are more than a minute apart or not. 
  mutate(total_time_mod=ifelse(t_hatdiff>=1,1,t_hat%%1))%>% #Record if more than a min apart or not
  #Deal with the first NA values for each subject
  mutate(total_time_mod=ifelse(is.na(total_time_mod),"missing",total_time_mod))%>%
  mutate(total_time_mod=as.numeric(ifelse(total_time_mod=="missing",
                               ifelse(lead(t_hat)-t_hat>=1,1,t_hat%%1),total_time_mod)
         ))%>%
  group_by(ID)%>%
  mutate(total_time_mod=ifelse(is.na(total_time_mod),t_hat%%1,total_time_mod))%>% #Some subjects only have 1 occurrence of light activity recorded so we need to account for their record
  summarise(total_time_mod=sum(total_time_mod))%>%
  as.data.frame()

##Let's count how many people have more than or equal to 150/7 min a day
mod_count_6<-mod_ac%>%
  dplyr::filter(total_time_mod>=150/7)
dim(mod_count_6) #Tells us how many people meet these recommendations which is 17.


##Total time in vig activity:
vig_ac<-min6_medday_reg$Y%>% #The registered output for visit 6
  rename(ID=id)%>% 
  transform(t_hat=t_hat*1440)%>%
  dplyr::filter(value>=5999)%>%
  group_by(ID)%>%
  mutate(t_hatdiff=t_hat-lag(t_hat))%>%#Find the difference between the minute recorded before which had mvig activity. This was done to see if the times are more than a minute apart or not. 
  mutate(total_time_mvig=ifelse(t_hatdiff>=1,1,t_hat%%1))%>% #Record if more than a min apart or not
  #Deal with the first NA values for each subject
  mutate(total_time_mvig=ifelse(is.na(total_time_mvig),"missing",total_time_mvig))%>%
  mutate(total_time_mvig=as.numeric(ifelse(total_time_mvig=="missing",
                               ifelse(lead(t_hat)-t_hat>=1,1,t_hat%%1),total_time_mvig)
         
         ))%>%
  group_by(ID)%>%
  mutate(total_time_mvig=ifelse(is.na(total_time_mvig),t_hat%%1,total_time_mvig))%>% #Some subjects only have 1 occurrence of light activity recorded so we need to account for their record
  summarise(total_time_vig=sum(total_time_mvig))%>%
  as.data.frame()

##Let's count how many people have more than or equal to 75/7 min a day
vig_count_6<-vig_ac%>%
  dplyr::filter(total_time_vig>=75/7)
dim(vig_count_6) #Tells us how many people meet these recommendations which is 4.

#Let's see if people are meeting both moderate and vig activity guidelines:
int<-inner_join(vig_count_6,mod_count_6,by="ID") #We find 1 person who meets these recommendations (ID 9589486 with vig time being  21 and mod time being 24.51288)

```
Repeat what I did above using the v8 information
```{r CDC Guidelines v8}
load("min8_medday_reg.Rdata")
#How many long are people doing moderate activity? 

mod_ac<-min8_medday_reg$Y%>% #The registered output for visit 6
  rename(ID=id)%>% 
  transform(t_hat=t_hat*1440)%>%
  dplyr::filter(between(value,2020,5998))%>%
  group_by(ID)%>%
  mutate(t_hatdiff=t_hat-lag(t_hat))%>%#Find the difference between the minute recorded before which had mod activity. This was done to see if the times are more than a minute apart or not. 
  mutate(total_time_mod=ifelse(t_hatdiff>=1,1,t_hat%%1))%>% #Record if more than a min apart or not
  #Deal with the first NA values for each subject
  mutate(total_time_mod=ifelse(is.na(total_time_mod),"missing",total_time_mod))%>%
  mutate(total_time_mod=as.numeric(ifelse(total_time_mod=="missing",
                               ifelse(lead(t_hat)-t_hat>=1,1,t_hat%%1),total_time_mod)
         ))%>%
  group_by(ID)%>%
  mutate(total_time_mod=ifelse(is.na(total_time_mod),t_hat%%1,total_time_mod))%>% #Some subjects only have 1 occurrence of light activity recorded so we need to account for their record
  summarise(total_time_mod=sum(total_time_mod))%>%
  as.data.frame()

##Let's count how many people have more than or equal to 150/7 min a day
mod_count<-mod_ac%>%
  dplyr::filter(total_time_mod>=150/7)
dim(mod_count) #Tells us how many people meet these recommendations which is 16.


##Total time in vig activity:
vig_ac<-min8_medday_reg$Y%>% #The registered output for visit 6
  rename(ID=id)%>% 
  transform(t_hat=t_hat*1440)%>%
  dplyr::filter(value>=5999)%>%
  group_by(ID)%>%
  mutate(t_hatdiff=t_hat-lag(t_hat))%>%#Find the difference between the minute recorded before which had mvig activity. This was done to see if the times are more than a minute apart or not. 
  mutate(total_time_mvig=ifelse(t_hatdiff>=1,1,t_hat%%1))%>% #Record if more than a min apart or not
  #Deal with the first NA values for each subject
  mutate(total_time_mvig=ifelse(is.na(total_time_mvig),"missing",total_time_mvig))%>%
  mutate(total_time_mvig=as.numeric(ifelse(total_time_mvig=="missing",
                               ifelse(lead(t_hat)-t_hat>=1,1,t_hat%%1),total_time_mvig)
         
         ))%>%
  group_by(ID)%>%
  mutate(total_time_mvig=ifelse(is.na(total_time_mvig),t_hat%%1,total_time_mvig))%>% #Some subjects only have 1 occurrence of light activity recorded so we need to account for their record
  summarise(total_time_vig=sum(total_time_mvig))%>%
  as.data.frame()

vig_count<-vig_ac%>%
  dplyr::filter(total_time_vig>=75/7)
dim(vig_count) #2 people

int<-inner_join(vig_count,mod_count,by="ID") #Nobody here meets both

##Let's see how many people consistently meet guidelines:
combo_mod<-inner_join(mod_count_6,mod_count,by="ID")
combo_vig<-inner_join(vig_count_6,vig_count,by="ID")

```
Generating predicted curves
```{r Median predicted activity visit6 and 8}
library(gghighlight)
library(grid)
load("min6_medday_reg.Rdata")
load("min8_medday_reg.Rdata")
#Predicted values--> Given by the registration function itself: 
preds<-data.frame(pred_value=(min6_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min6_medday_reg$Y,preds)

#Replace ids with numbers
id_rep<-data.frame(id=visits_ID$ID,p_id=c(1:1343))
combo_rep<-inner_join(combo,id_rep,by="id")

##Graph of the predicted activity curves at visit 6
post_reg_gaus_v6<-ggplot(combo_rep,aes(x=(t_hat*1440),y=pred_value,col=p_id,group=p_id))+geom_line() +labs(title="48-Month",x = "Registered time (Minute)", y = "Predicted Activity Count",
              tag = "A") +labs(color = "Participant ID")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 30))+scale_color_viridis_c()

##Graph of the predicted activity curves at visit 8
preds<-data.frame(pred_value=(min8_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min8_medday_reg$Y,preds)

post_reg_gaus_v8<-ggplot(combo_rep,aes(x=(t_hat*1440),y=pred_value,col=p_id,group=p_id))+geom_line() +labs(title="72-Month",x = "Registered time (Minute)", y = "Predicted Activity Count",tag = "B") +labs(color = "Participant ID")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 35))+scale_color_viridis_c()

##Plot side-by-side
plot_grid(post_reg_gaus_v6,post_reg_gaus_v8,ncol = 1, nrow = 2)

```

```{r ARUC variety of methods}
library(bayestestR)
load("min6_medday_reg.Rdata")
load("min8_medday_reg.Rdata")
setwd(("/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac"))

#Visit 6:
preds<-data.frame(pred_value=(min6_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min6_medday_reg$Y,preds)

combo_test_v6<-combo%>%
  group_by(id)%>%
  rename(ID=id)%>%
  mutate(area_trap=area_under_curve(t_hat, pred_value, method = "trapezoid"))%>%
  mutate(area_step=area_under_curve(t_hat, pred_value, method = "step"))%>%
  mutate(area_spline=area_under_curve(t_hat, pred_value, method = "spline"))%>%
  summarise_all(last)%>% #Get back the area under the curve for each subject -last val
  as.data.frame()%>%
  rename(AURC=area_trap)%>%
  select(ID,AURC)

save(combo_test_v6,file="/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/combo_test_v6.Rdata")


#Visit 8:
preds<-data.frame(pred_value=(min8_medday_reg$fpca_obj$Yhat$value))
combo<-cbind(min8_medday_reg$Y,preds)

combo_test_v8<-combo%>%
  group_by(id)%>%
  rename(ID=id)%>%
  mutate(area_trap=area_under_curve(t_hat, pred_value, method = "trapezoid"))%>%
  mutate(area_step=area_under_curve(t_hat, pred_value, method = "step"))%>%
  mutate(area_spline=area_under_curve(t_hat, pred_value, method = "spline"))%>%
  summarise_all(last)%>% #Get back the area under the curve for each subject -last val
  as.data.frame()%>%
  rename(AURC=area_trap)%>%
  select(ID,AURC)

save(combo_test_v8,file="/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/combo_test_v8.Rdata")

```


PRIMARY Analysis (OA outcomes)

Looking at demographics
```{r AURC across sex}
load("combo_test_v6.Rdata") #AURC data
load("combo_test_v8.Rdata") #AURC data

#First we remove those in the control group:
load("visits_ID.Rdata")
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")

no_control_ID<-OAI_KL_JSW_data%>%
  dplyr::filter(V00COHORT!="3: Non-exposed control group")%>%
  dplyr::filter(VISIT=="V00")%>%#we only need control IDs
  select(ID)%>%
  inner_join(visits_ID,by="ID")#sample pop excluding control group

demo_v6<-OAI_KL_JSW_data%>%
  select(ID,SEX,BMI,RACE,AGE,V00COHORT,VISIT)%>%
  dplyr::filter(VISIT=="V06")%>%
  dplyr::filter(ID %in% no_control_ID$ID)
demo_v6<-inner_join(demo_v6,combo_test_v6,by="ID")

demo_v8<-OAI_KL_JSW_data%>%
  select(ID,SEX,BMI,RACE,AGE,V00COHORT,VISIT)%>%
  dplyr::filter(VISIT=="V08")%>%
  dplyr::filter(ID %in% no_control_ID$ID)
demo_v8<-inner_join(demo_v8,combo_test_v8,by="ID")

combined_data<-rbind(demo_v6,demo_v8) #PUT AURC data together

# Filter data for men and run Wilcoxon Signed-Rank Test
men_data<-combined_data%>%
  dplyr::filter(SEX=="M")
wilcox_test_men <- wilcox.test(AURC ~ VISIT, data = men_data, paired = TRUE)
print(wilcox_test_men)

# Filter data for women and run Wilcoxon Signed-Rank Test
women_data<-combined_data%>%
  dplyr::filter(SEX=="F")
wilcox_test_women <- wilcox.test(AURC ~ VISIT, data = women_data, paired = TRUE)
print(wilcox_test_women)

#Perform Mann-Whitney U test (unpaired) at v6
v6men_data<-combined_data%>%
  dplyr::filter(SEX=="M")%>%
  dplyr::filter(VISIT=="V06")

v6women_data<-combined_data%>%
  dplyr::filter(SEX=="F")%>%
  dplyr::filter(VISIT=="V06")

result <- wilcox.test(v6men_data$AURC, v6women_data$AURC)
print(result)

#Perform Mann-Whitney U test (unpaired) at v8
v8men_data<-combined_data%>%
  dplyr::filter(SEX=="M")%>%
  dplyr::filter(VISIT=="V08")

v8women_data<-combined_data%>%
  dplyr::filter(SEX=="F")%>%
  dplyr::filter(VISIT=="V08")

result <- wilcox.test(v8men_data$AURC, v8women_data$AURC)
```

```{r AURC across race}
#Using demo data from the previous chunk

v6white_dem<-demo_v6%>%
  dplyr::filter(RACE=="White")

v8white_dem<-demo_v8%>%
  dplyr::filter(RACE=="White")

v6nw_dem<-demo_v6%>%
  dplyr::filter(RACE!="White")

v8nw_dem<-demo_v8%>%
  dplyr::filter(RACE!="White")

#Perform Mann-Whitney U test (unpaired) at v6
result <- wilcox.test(v6white_dem$AURC, v6nw_dem$AURC)
#Perform Mann-Whitney U test (unpaired) at v8
result <- wilcox.test(v8white_dem$AURC, v8nw_dem$AURC)

#Combine data first
dem<-rbind(demo_v6,demo_v8)


# Filter data for white and run Wilcoxon Signed-Rank Test
white_dem<-dem%>%
  dplyr::filter(RACE=="White")

wilcox_test_white <- wilcox.test(AURC ~ VISIT, data = white_dem, paired = TRUE)
print(wilcox_test_white)


# Filter data for nonwhite and run Wilcoxon Signed-Rank Test
nw_dem<-dem%>%
  dplyr::filter(RACE!="White")

wilcox_test_nw <- wilcox.test(AURC ~ VISIT, data = nw_dem, paired = TRUE)
print(wilcox_test_nw)

```

```{r AURC across age}
#IDS of those above or below 64 at 48-month
above_dem<-demo_v6%>%
  dplyr::filter(AGE>=64)%>%
  select(ID)

below_dem<-demo_v6%>%
  dplyr::filter(AGE<64)%>%
  select(ID)

above_v6_dem<-demo_v6%>%
  dplyr::filter(ID %in% above_dem$ID)
below_v6_dem<-demo_v6%>%
  dplyr::filter(ID %in% below_dem$ID)
#Perform Mann-Whitney U test (unpaired) at v6
result <- wilcox.test(above_v6_dem$AURC, below_v6_dem$AURC)

#Visit8
above_v8_dem<-demo_v8%>%
  dplyr::filter(ID %in% above_dem$ID)
below_v8_dem<-demo_v8%>%
  dplyr::filter(ID %in% below_dem$ID)
#Perform Mann-Whitney U test (unpaired) at v8
result <- wilcox.test(above_v8_dem$AURC, below_v8_dem$AURC)

#Paired test
dem<-rbind(demo_v6,demo_v8)

above_age_dem<-dem%>%
  dplyr::filter(ID %in% above_dem$ID)
wilcox_test_above <- wilcox.test(AURC ~ VISIT, data = above_age_dem, paired = TRUE)
print(wilcox_test_above)

below_age_dem<-dem%>%
  dplyr::filter(ID %in% below_dem$ID)
wilcox_test_below <- wilcox.test(AURC ~ VISIT, data = below_age_dem, paired = TRUE)
print(wilcox_test_below)
```

```{r AURC subcohort}
##Progression
prog_demo_v6<-demo_v6%>%
  dplyr::filter(V00COHORT=="1: Progression")
prog_demo_v8<-demo_v8%>%
  dplyr::filter(V00COHORT=="1: Progression")

#Paired test
prog_dem<-rbind(prog_demo_v6,prog_demo_v8)

wilcox_test_prog <- wilcox.test(AURC ~ VISIT, data = prog_dem, paired = TRUE)
print(wilcox_test_prog)

##Repeat for incidence
inc_demo_v6<-demo_v6%>%
  dplyr::filter(V00COHORT=="2: Incidence")
inc_demo_v8<-demo_v8%>%
  dplyr::filter(V00COHORT=="2: Incidence")

#Paired test
inc_dem<-rbind(inc_demo_v6,inc_demo_v8)

wilcox_test_inc <- wilcox.test(AURC ~ VISIT, data = inc_dem, paired = TRUE)
print(wilcox_test_inc)

#Repeat for control:
con_demo_v6<-demo_v6%>%
  dplyr::filter(V00COHORT=="3: Non-exposed control group")
con_demo_v8<-demo_v8%>%
  dplyr::filter(V00COHORT=="3: Non-exposed control group")

#Paired test
con_dem<-rbind(con_demo_v6,con_demo_v8)

wilcox_test_con <- wilcox.test(AURC ~ VISIT, data = con_dem, paired = TRUE)
print(wilcox_test_con)

combo_v6<-rbind(inc_demo_v6,prog_demo_v6)
combo_v8<-rbind(inc_demo_v8,prog_demo_v8)
# Run the Kruskal-Wallis test v6
result_v6 <- wilcox.test(AURC ~ V00COHORT, data = combo_v6)

# Run the Kruskal-Wallis test v8
result_v8 <- wilcox.test(AURC ~ V00COHORT, data = combo_v8)
print(result_v8)
```

```{r AURC across BMI}
load("demo_v6.Rdata")
load("demo_v8.Rdata")

demo_v6BMI<-demo_v6%>%
  mutate(BMI = case_when(BMI <= 24.999 ~ 'Healthy weight',
  between(BMI, 25, 30)~ 'Overweight',
  BMI >30 ~ 'Obese'))

#Underweight
#underweight_v6<-demo_v6BMI%>%
  #dplyr::filter(BMI=="Underweight")
#Healthy weight
hweight_v6<-demo_v6BMI%>%
  dplyr::filter(BMI=="Healthy weight")
#Overweight
ovweight_v6<-demo_v6BMI%>%
  dplyr::filter(BMI=="Overweight")
#Obese
obweight_v6<-demo_v6BMI%>%
  dplyr::filter(BMI=="Obese")

###V8
#underweight_v8<-demo_v8%>%
 # dplyr::filter(ID %in%underweight_v6$ID)%>%
 # mutate(BMI="Underweight")
hweight_v8<-demo_v8%>%
  dplyr::filter(ID %in%hweight_v6$ID)%>%
  mutate(BMI="Healthy weight")
ovweight_v8<-demo_v8%>%
  dplyr::filter(ID %in%ovweight_v6$ID)%>%
  mutate(BMI="Overweight")
obweight_v8<-demo_v8%>%
  dplyr::filter(ID %in%obweight_v6$ID)%>%
  mutate(BMI="Obese")

##Running Paired tests:
#Underweight 
underweight_dem<-rbind(underweight_v6,underweight_v8)

wilcox_test_und <- wilcox.test(AURC ~ VISIT, data = underweight_dem, paired = TRUE)
print(wilcox_test_und)

#Healthy weight
hweight_dem<-rbind(hweight_v6,hweight_v8)

wilcox_test_hw <- wilcox.test(AURC ~ VISIT, data = hweight_dem, paired = TRUE)
print(wilcox_test_hw)

#Overweight
ovweight_dem<-rbind(ovweight_v6,ovweight_v8)

wilcox_test_ow <- wilcox.test(AURC ~ VISIT, data = ovweight_dem, paired = TRUE)
print(wilcox_test_ow)

#Obese
obweight_dem<-rbind(obweight_v6,obweight_v8)

wilcox_test_ob <- wilcox.test(AURC ~ VISIT, data = obweight_dem, paired = TRUE)
print(wilcox_test_ob)

##Running unpaired tests:
#V6
v6_BMI<-rbind(hweight_v6,ovweight_v6,obweight_v6)
kruskal_result_v6 <- kruskal.test(AURC ~ BMI, data = v6_BMI)
print(kruskal_result_v6)

#V8
v8_BMI<-rbind(hweight_v8,ovweight_v8,obweight_v8)
kruskal_result_v8 <- kruskal.test(AURC ~ BMI, data = v8_BMI)
print(kruskal_result_v8)

```

WOMAC 
```{r Linear mixed effects model}
library(mmrm)
library(dplyr)
setwd("/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac")
load("combo_test_v6.Rdata")
load("combo_test_v8.Rdata")
load("visits_ID.Rdata")
load("Outcomes.Rdata")#Called OAI_KL_JSW_data

area_v8<-combo_test_v8%>%
  transform(ID=as.factor(ID),
            area=AURC)%>%
  as.data.frame()

area_v6<-combo_test_v6%>%
  transform(ID=as.factor(ID),
            area=AURC)%>%
  as.data.frame()

#Extract the valid information form each visit:
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE,WOMTSmax)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()

#By doing it visit by visit I make sure I get those with valid info at both visits. 
outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE,WOMTSmax)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()#Removes missing ppl with invalid values

##The above are subjects in the sample population with valid BMI and WOMAC at each visit to find those who have BMI values at both visits:
ids<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1296 have BMI and WOMAC at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 1296
  dplyr::filter(ID %in%ids)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids) #Filter the ARUC data to include only those 1296

#Combine the areas with the outcomes data above
v6_comb<-inner_join(outcomes_v6,area_v6,by="ID")
v8_comb<-inner_join(outcomes_v8,area_v8,by="ID")

combo_visits<-rbind(v6_comb,v8_comb)
combo_visits<-combo_visits%>%
  dplyr::filter(ID %in% visits_ID$ID)

##Fit model:
womac_results <- mmrm(formula = WOMTSmax ~ RACE+SEX+AGE+BMI+area+us(VISIT|ID),
  data = combo_visits)
#I saved these results:
save(womac_results,file="/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/womac_results.Rdata")
```

KL model
```{r}
KL_area_v6<-combo_test_v6%>%
  transform(ID=as.factor(ID))%>%
  rename(area=AURC)%>%
  as.data.frame()

KL_area_v8<-combo_test_v8%>%
  transform(ID=as.factor(ID))%>%
  rename(area=AURC)%>%
  as.data.frame()

#Extract the valid information form each visit:
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE,XRKL)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()

#By doing it visit by visit I make sure I get those with valid info at both visits. 
outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE,XRKL)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()#Removes missing ppl with invalid values

##The above are subjects in the sample population with valid BMI and KL at each visit to find those who have BMI values at both visits:
ids<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #886 have BMI and KL at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 886
  dplyr::filter(ID %in%ids)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids) #Filter the ARUC data to include only those 886

#Combine the areas with the outcomes data above
v6_comb<-inner_join(outcomes_v6,KL_area_v6,by="ID")
v8_comb<-inner_join(outcomes_v8,KL_area_v8,by="ID")

combo_visits_KL<-rbind(v6_comb,v8_comb)

##Change the visit to be numerical:
KL_data_pop<-combo_visits_KL%>%
  mutate(VISIT = recode(VISIT, 'V06' = 48, 'V08' = 72))%>%
  data.frame()

cr_vals <- cr_setup(KL_data_pop$XRKL,direction = "backward") 
cr_data <- KL_data_pop[cr_vals$subs, ]
cr_data$XRKL_new <- cr_vals$y ## Note that this needs to be y, not XRKL
cr_data$cohort <- cr_vals$cohort
cr_data$VISIT <- factor(cr_data$VISIT)
cr_data$cohort <- relevel(cr_data$cohort, ref = 4) #Reference level changed

KL_results <- mixed_model(XRKL_new ~ cohort + RACE+SEX+AGE+BMI+area, random = ~ VISIT||ID, 
                          data = cr_data, family = binomial(), control = list(max_coef_value=100, iter_EM = 0))
```

Supplementary Analysis: (Found in the supplementary analysis)
```{r Sup: No control Analysis}
#We are running no control registration 
setwd(("/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac"))
load("min6.Rdata")#Cleaned min data
load("min8.Rdata")#Cleaned min data
load("visits_ID.Rdata")
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")
library(gghighlight)
library(grid)

no_control_ID<-OAI_KL_JSW_data%>%
  dplyr::filter(V00COHORT=="3: Non-exposed control group")%>%
  dplyr::filter(VISIT=="V00")%>%#we only need control IDs
  select(ID)%>%
  inner_join(visits_ID,by="ID")#sample pop in control group


min6_medday<-min6%>%
  group_by(ID,V06MinSequence)%>%
  summarise(value=median(V06MINCnt))%>%
  rename(id=ID)%>%
  rename(index=V06MinSequence)%>%
  dplyr::filter(!id %in%no_control_ID$ID)%>%
  dplyr::filter(id%in% visits_ID$ID)%>%
  na.omit()%>%
  as.data.frame()##This must be done to run registration

min8_medday<-min8%>%
  group_by(ID,V08MinSequence)%>%
  summarise(value=median(V08MINCnt))%>%
  rename(id=ID)%>%
  rename(index=V08MinSequence)%>%
  dplyr::filter(!id %in% no_control_ID$ID)%>%
  dplyr::filter(id%in% visits_ID$ID)%>%
  na.omit(min8_medday)%>%
  as.data.frame()##This must be done to run registration

#This creates median days excluding those in the control group
#Visit 6
registr_gauss2_v6_noc<-register_fpca(Y =min6_medday, family = "gaussian",
                    npc = 3, 
                    Kt=8,
                    Kh=4,
                    gradient = FALSE,
                    warping = "piecewise_linear2")
#I saved these results:
save(registr_gauss2_v6_noc,file="registr_gauss_v6_noc.Rdata")

#Visit 8
registr_gauss2_v8_noc<-register_fpca(Y =min8_medday, family = "gaussian",
                    npc = 3, 
                    Kt=8,
                    Kh=4,
                    gradient = FALSE,
                    warping = "piecewise_linear2")

#I saved these results:
save(registr_gauss2_v8_noc,file="/gpfs/fs1/home/o/oespinga/reshani0/registr_gauss2_v8_noc.Rdata")

##Make the plots:
setwd("/gpfs/fs0/scratch/o/oespinga/reshani0/")

##No control:
exc_control_ID<-OAI_KL_JSW_data%>%
  dplyr::filter(V00COHORT!="3: Non-exposed control group")%>%
  dplyr::filter(VISIT=="V00")%>%#we only need control IDs
  select(ID)%>%
  inner_join(visits_ID,by="ID")#sample pop in control group

##Visit 6:
preds<-data.frame(pred_value=(registr_gauss2_v6_noc$fpca_obj$Yhat$value))
combo<-cbind(registr_gauss2_v6_noc$Y,preds)

#Replace ids with numbers
id_rep<-data.frame(id=exc_control_ID$ID,p_id=c(1:1337))
combo_rep<-inner_join(combo,id_rep,by="id")

pre_6reg<-ggplot(combo_rep,aes(x=(tstar),y=value,col=p_id,group=p_id))+geom_line(show.legend = FALSE)+labs(x = "Unregistered time (Minute)", y = "Median activity count",tag = "A")+labs(color = "Participant Id")+
  scale_x_continuous(breaks=seq(0, 1440, 120))+theme(text = element_text(size = 8))+theme(axis.title=element_text(size=8))+ scale_color_viridis_c()+ggtitle("48-month visit")+theme(plot.title = element_text(size=10,face="bold"))+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

#Warping
warping_6p_gaus<-ggplot(combo_rep, aes(x = tstar, y = (t_hat*1440), col = (p_id)))+
    geom_line(show.legend = FALSE) + labs(x = "Unregistered time (Minute)", y = "Registered time (Minute)",
              tag = "C")+labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_y_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 20))+theme(text = element_text(size = 10))+theme(axis.title=element_text(size=10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

##Post registration:
post_reg_gaus_6p<-ggplot(combo_rep,aes(x=(t_hat*1440),y=value,col=(p_id),group=p_id))+
    geom_line(show.legend = FALSE) +labs(x = "Registered time (Minute)", y = "Median Activity Count",tag = "E") +labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 10))+theme(text = element_text(size = 10))+theme(axis.title=element_text(size=10))+scale_color_viridis_c()+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))


##Visit 8:
preds<-data.frame(pred_value=(registr_gauss2_v8_noc$fpca_obj$Yhat$value))
combo<-cbind(registr_gauss2_v8_noc$Y,preds)

#Replace ids with numbers
id_rep<-data.frame(id=exc_control_ID$ID,p_id=c(1:1337))
combo_rep<-inner_join(combo,id_rep,by="id")


#Pre-registration curve
pre_8reg<-ggplot(combo_rep,aes(x=(tstar),y=value,col=p_id,group=p_id))+
    geom_line(show.legend = FALSE)+ labs(x = "Unregistered time (Minute)", y = "Median activity count",
              tag = "B")+labs(color = "Subject Id")+
  scale_color_gradient(low="blue", high="grey")+ scale_x_continuous(breaks=seq(0, 1440, 120))+theme(text = element_text(size = 20))+theme(text = element_text(size = 10))+theme(text = element_text(size = 10))+scale_color_viridis_c()+theme(axis.title=element_text(size=10))+ggtitle("72-month visit")+scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(axis.title=element_text(size=8))+theme(plot.title = element_text(size=10,face="bold"))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

#Warping
warping_8p_gaus<-ggplot(combo_rep, aes(x = tstar, y = (t_hat*1440), col = (id)))+
    geom_line(show.legend = FALSE)+ labs( x = "Unregistered time (Minute)", y = "Registered time (Minute)",
              tag = "D")+labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_y_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size = 10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size = 0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

##Post registration:
post_reg_gaus_8p<-ggplot(combo_rep,aes(x=(t_hat*1440),y=value,group=p_id,col=(p_id)))+
    geom_line(show.legend = FALSE) +labs(x = "Registered time (Minute)", y = "Median Activity Count",
              tag = "F") +labs(color = "Subject Id")+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_x_continuous(breaks=seq(0, 1440, 120))+scale_color_gradient(low="blue", high="grey")+theme(text = element_text(size =10))+scale_color_viridis_c()+theme(axis.title=element_text(size=8))+ scale_y_continuous(breaks=seq(0, 8000, 2000))+theme(text = element_text(size = 8))+ geom_line(aes(color = p_id, group = p_id), size =0.5,color="red",
            data = ~ subset(., p_id %in% c(168)))

mix_plotnoc<-grid.arrange(pre_6reg,pre_8reg,
                  warping_6p_gaus, warping_8p_gaus,
                  post_reg_gaus_6p,post_reg_gaus_8p,
                  ncol = 2, nrow = 3)


ggsave(plot= mix_plotnoc, filename = '/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/noccomboplot.pdf',width = 297,
  height = 210,
  units = c( "mm"))








##I will run the primary analysis
##Find AURC
library(bayestestR)

#Visit 6:
preds<-data.frame(pred_value=(registr_gauss2_v6_noc$fpca_obj$Yhat$value))
combo<-cbind(registr_gauss2_v6_noc$Y,preds)

combo_test_v6_noc<-combo%>%
  group_by(id)%>%
  rename(ID=id)%>%
  mutate(area_trap=area_under_curve(t_hat, pred_value, method = "trapezoid"))%>%
  summarise_all(last)%>% #Get back the area under the curve for each subject -last val
  as.data.frame()

#Repeat for visit 8
preds<-data.frame(pred_value=(registr_gauss2_v8_noc$fpca_obj$Yhat$value))
combo<-cbind(registr_gauss2_v8_noc$Y,preds)

combo_test_v8_noc<-combo%>%
  group_by(id)%>%
  rename(ID=id)%>%
  mutate(area_trap=area_under_curve(t_hat, pred_value, method = "trapezoid"))%>%
  summarise_all(last)%>% #Get back the area under the curve for each subject -last val
  as.data.frame()

save(combo_test_v8_noc,file="/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac/combo_test_v8_noc.Rdata")

library(mmrm)
library(tidyverse)
setwd("/gpfs/fs1/home/o/oespinga/reshani0/reshani_prac")
load("combo_test_v6.Rdata")
load("combo_test_v8.Rdata")
load("visits_ID.Rdata")
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")

area_v8<-combo_test_v8_noc%>%
  transform(ID=as.factor(ID))%>%
  as.data.frame()

area_v6<-combo_test_v6_noc%>%
  transform(ID=as.factor(ID))%>%
  as.data.frame()

#Extract the valid information form each visit:
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% no_control_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE,WOMTSmax)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()

#By doing it visit by visit I make sure I get those with valid info at both visits. 
outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% no_control_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE,WOMTSmax)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()#Removes missing ppl with invalid values

##The above are subjects in the sample population with valid BMI and WOMAC at each visit to find those who have BMI values at both visits:
ids<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1290 have BMI and WOMAC at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 1290
  dplyr::filter(ID %in%ids)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids) #Filter the ARUC data to include only those 1290

#Combine the areas with the outcomes data above
v6_comb<-inner_join(outcomes_v6,area_v6,by="ID")
v8_comb<-inner_join(outcomes_v8,area_v8,by="ID")

combo_visits<-rbind(v6_comb,v8_comb)
combo_visits<-combo_visits%>%
  dplyr::filter(ID %in% no_control_ID$ID)

##Fit model:
womac_results <- mmrm(formula = WOMTSmax ~ RACE+SEX+AGE+BMI+area_trap+us(VISIT|ID), data = combo_visits) #Excluding control group

###KL 
KL_area_v6<-combo_test_v6_noc%>%
  transform(ID=as.factor(ID))%>%
  as.data.frame()

KL_area_v8<-combo_test_v8_noc%>%
  transform(ID=as.factor(ID))%>%
  as.data.frame()

#Extract the valid information form each visit:
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% no_control_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE,XRKL)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()

#By doing it visit by visit I make sure I get those with valid info at both visits. 
outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% no_control_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE,XRKL)%>%
  transform(VISIT=as.factor(VISIT))%>%
  transform(ID=as.factor(ID))%>%
  na.omit()#Removes missing ppl with invalid values

##The above are subjects in the sample population with valid BMI and KL at each visit to find those who have BMI values at both visits:
ids<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #881 have BMI and KL at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 886
  dplyr::filter(ID %in%ids)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids) #Filter the ARUC data to include only those 886

#Combine the areas with the outcomes data above
v6_comb<-inner_join(outcomes_v6,KL_area_v6,by="ID")
v8_comb<-inner_join(outcomes_v8,KL_area_v8,by="ID")

combo_visits_KL<-rbind(v6_comb,v8_comb)

##Change the visit to be numerical:
KL_data_pop<-combo_visits_KL%>%
  mutate(VISIT = recode(VISIT, 'V06' = 48, 'V08' = 72))%>%
  data.frame()

cr_vals <- cr_setup(KL_data_pop$XRKL)
cr_data <- KL_data_pop[cr_vals$subs, ]
cr_data$XRKL_new <- cr_vals$y ## Note that this needs to be y, not XRKL
cr_data$cohort <- cr_vals$cohort
cr_data$VISIT <- factor(cr_data$VISIT)
cr_data$cohort <- relevel(cr_data$cohort, ref = 4)

KL_results <- mixed_model(XRKL_new ~ cohort + RACE+SEX+AGE+BMI+area_trap, random = ~ VISIT|ID, 
                  data = cr_data, family = binomial(), control = list(max_coef_value=100, iter_EM = 0))
```

```{r Sup: WOMAC pain and stiffness}
womac_pain_6<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/AllClinical06.txt", sep="|", header=T)
womac_pain_8<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/AllClinical08.txt", sep="|", header=T)
load("visits_ID.Rdata")

womac_6<-womac_pain_6%>%
  select(ID,V06WOMKPL,V06WOMKPR,V06WOMSTFL,V06WOMSTFR)%>%
  dplyr::filter(ID %in% visits_ID$ID)%>%
  mutate(max_WOMKP_6=pmax(V06WOMKPL,V06WOMKPR,na.rm=TRUE), #Took max across the two visits
         max_WOMSTF_6=pmax(V06WOMSTFL,V06WOMSTFR,na.rm=TRUE))%>%
  as.data.frame()

womac_pain_8<-womac_pain_8%>%
  select(ID,V08WOMKPL,V08WOMKPR,V08WOMSTFL,V08WOMSTFR)%>%
  dplyr::filter(ID %in% visits_ID$ID)%>%
  mutate(max_WOMKP_8=pmax(V08WOMKPL,V08WOMKPR,na.rm=TRUE), #Took max across the two visits
         max_WOMSTF_8=pmax(V08WOMSTFL,V08WOMSTFR,na.rm=TRUE))%>%
  as.data.frame()

## NA cases there exists an NA for V06WOMSTFL, and 2 for V08WOMSTFL dealt by using the other value provided.
#Now we combine the above:
womac_visits<-inner_join(womac_6,womac_pain_8,by="ID")%>%
  select(ID,max_WOMKP_6,max_WOMSTF_6,max_WOMKP_8,max_WOMSTF_8)%>%
  as.data.frame() #No missing values

##Run the same model we ran for WOMAC score in primary:
load("combo_test_v6.Rdata")#File with AURC for each subject on visit 8 on median day
load("combo_test_v8.Rdata")#File with AURC for each subject on visit 6 on median day
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")#This is the outcome data

##Working with outcomes data 
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()
#By doing it visit by visit I make sure I get those with valid BMI at both visits. 

outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()

##The above are subjects in the sample population with valid BMI at each visit to find those who have BMI values at both visits:
ids_BMI<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1309 have BMI at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 1309
  dplyr::filter(ID %in%ids_BMI)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids_BMI) #Filter the AURC data to include only those 1309

##Creating data frame for model
#V6
womac_v6<-womac_visits%>%
  select(ID,max_WOMKP_6,max_WOMSTF_6)%>%
  rename(WOMKP=max_WOMKP_6,
         WOMSTF=max_WOMSTF_6)%>%
  inner_join(combo_test_v6, by="ID")%>%
  as.data.frame()%>%
  inner_join(outcomes_v6,by="ID")%>%
  transform(ID=as.factor(ID))

womac_v8<-womac_visits%>%
  select(ID,max_WOMKP_8,max_WOMSTF_8)%>%
  rename(WOMKP=max_WOMKP_8,
         WOMSTF=max_WOMSTF_8)%>%
  inner_join(combo_test_v8, by="ID")%>%
  as.data.frame()%>%
  inner_join(outcomes_v8,by="ID")%>%
  transform(ID=as.factor(ID))#Making ID as factor part of the mmrm model fit.

womac_combo<-rbind(womac_v6,womac_v8)

#Using pain
pain_fit <- mmrm(formula = WOMKP ~ RACE+SEX+AGE+BMI+AURC+us(VISIT|ID),
  data = womac_combo)
summary(pain_fit)

#Using Stiffness
stiff_fit <- mmrm(formula = WOMSTF ~ RACE+SEX+AGE+BMI+AURC+us(VISIT|ID),
  data = womac_combo)
summary(stiff_fit)

```



```{r Sup: JSW}
qjsw_6<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_qjsw_duryea06.txt", sep="|", header=T)
qjsw_8<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_qjsw_duryea08.txt", sep="|", header=T)
load("visits_ID.Rdata")

qjsw_6<-qjsw_6%>%
  select(ID,side,V06MCMJSW,V06BARCDJD)%>%
  group_by(ID)%>%
  dplyr::filter(ID %in%visits_ID$ID)%>%
  na.omit()%>% #Remove NA values measured for JSW 
  group_by(ID,side)%>%
  dplyr::summarise(V06MCMJSW=max(V06MCMJSW))%>% #Finds the max measured for each side with repeats
  as.data.frame()%>%
  group_by(ID)%>%
  dplyr::summarise(jsw_comp=sum(V06MCMJSW))

qjsw_8<-qjsw_8%>%
  select(ID,side,V08MCMJSW,V08BARCDJD)%>%
  group_by(ID)%>%
  dplyr::filter(ID %in%visits_ID$ID)%>%
  na.omit()%>% #Remove NA values measured for JSW 
  group_by(ID,side)%>%
  dplyr::summarise(V08MCMJSW=max(V08MCMJSW))%>% #Finds the max measured for each side with repeats
  as.data.frame()%>% #Leaving us with 823 subjects 
  group_by(ID)%>%
  dplyr::summarise(jsw_comp=sum(V08MCMJSW))


##Run the same model we ran for WOMAC score in primary:
load("combo_test_v6.Rdata")#File with AURC for each subject on visit 8 on median day
load("combo_test_v8.Rdata")#File with AURC for each subject on visit 6 on median day
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")#This is 

##Working with outcomes data 
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()
#By doing it visit by visit I make sure I get those with valid BMI at both visits. 

outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()

##The above are subjects in the sample population with valid BMI at each visit to find those who have BMI values at both visits:
ids_BMI<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1309 have BMI at both visits.

outcomes_v6<-outcomes_v6%>% #Filter the ARUC data to include only those 1309
  dplyr::filter(ID %in%ids_BMI)
outcomes_v8<-outcomes_v8%>%
  dplyr::filter(ID %in%ids_BMI) #Filter the AURC data to include only those 1309


##Creating data frame for model
#V6
jsw_v6<-qjsw_6%>%
  inner_join(combo_test_v6, by="ID")%>%
  as.data.frame()%>%
  inner_join(outcomes_v6,by="ID")%>%
  transform(ID=as.factor(ID))

jsw_v8<-qjsw_8%>%
  inner_join(combo_test_v8, by="ID")%>%
  as.data.frame()%>%
  inner_join(outcomes_v8,by="ID")%>%
  transform(ID=as.factor(ID))%>%
  dplyr::filter(ID %in% jsw_v6$ID) #We only want those present at both visits

#We only want those present at both visits
jsw_v6<-jsw_v6%>%
  dplyr::filter(ID %in% jsw_v8$ID)

jsw_combo<-rbind(jsw_v6,jsw_v8)

#Using JSW
jsw_fit <- mmrm(formula = jsw_comp ~ RACE+SEX+AGE+BMI+AURC+us(VISIT|ID),
  data = jsw_combo)
summary(jsw_fit)

```

```{r Sup: Chondrocalcinosis}
osteo_6<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_sq_bu06.txt", sep="|", header=T)
osteo_8<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_sq_bu08.txt", sep="|", header=T)
load("visits_ID.Rdata")#IDs of those in the sample pop 

#We will make the outcome binary. This means if at least one has a score not 0 across the medial or lateral sides they get scored 1 if 0 across all then 0.
osteoph_6_chond<-osteo_6%>%
  dplyr::filter(ID %in% visits_ID$ID)%>% #This left us with 1309 subjects I had 1309 unique barcodes
  select(ID,SIDE,V06XRCHM,V06XRCHL,V06BARCDBU)%>%
  na.omit()%>% #Remove NA values measured for medial 
  group_by(ID,SIDE)%>%#This leaves us 823 subjects
  dplyr::summarise(V06XRCHM=max(V06XRCHM),
                   V06XRCHL=max(V06XRCHL))%>%
  as.data.frame()%>% #Leaving us with 823 subjects 
  group_by(ID)%>%
  dplyr::summarise(medial_comp=sum(V06XRCHM),
                   lateral_comp=sum(V06XRCHL))%>%
  mutate(medial_comp=ifelse(medial_comp>=1,1,0),
         lateral_comp=ifelse(lateral_comp>=1,1,0))%>%
  mutate(chond=ifelse(medial_comp==0&lateral_comp==0,0,1))%>%#Making the outcome into a binary where those with 1 on at least one side is graded as a 1 and all else is a 0.
  mutate(VISIT=as.factor(c("V06"))) #Leaves us with 823 ppl

osteoph_8_chond<-osteo_8%>%
  dplyr::filter(ID %in% visits_ID$ID)%>% #This left us with 1309 subjects I had 1309 unique barcodes
  select(ID,SIDE,V08XRCHM,V08XRCHL,V08BARCDBU)%>%
  na.omit()%>% #Remove NA values measured for medial 
  group_by(ID,SIDE)%>%#This leaves us 823 subjects
  dplyr::summarise(V08XRCHM=max(V08XRCHM),
                   V08XRCHL=max(V08XRCHL))%>%
  as.data.frame()%>% #Leaving us with 823 subjects 
  group_by(ID)%>%
  dplyr::summarise(medial_comp=sum(V08XRCHM),
                   lateral_comp=sum(V08XRCHL))%>%
  mutate(medial_comp=ifelse(medial_comp>=1,1,0),
         lateral_comp=ifelse(lateral_comp>=1,1,0))%>%
  mutate(chond=ifelse(medial_comp==0&lateral_comp==0,0,1))%>%#Making the outcome into a binary where those with 1 on at least one side is graded as a 1 and all else is a 0.
  mutate(VISIT=as.factor(c("V08")))

#I want to check if I have the same IDs at both visits
identical(osteoph_8_chond$ID,osteoph_6_chond$ID) #I get back false

#Get the IDs of people with measurements at both visits 
#Extract list of IDs from each dataset and find shared ID list. 
chond_6_ids<-osteoph_6_chond%>%
  select(ID)

chond_8_ids<-osteoph_8_chond%>%
  select(ID)

ids_chond<-inner_join(chond_6_ids,chond_8_ids,by="ID") #Leaving me with 115

##Combine the outcomes for each visit
chond_outcome<-rbind(osteoph_8_chond,osteoph_6_chond)%>%
  dplyr::filter(ID %in% ids_chond$ID)%>%
  group_by(ID)%>%
  as.data.frame()##Final with medial outcomes

load("combo_test_v6.Rdata")#File with AURC for each subject on visit 8 on median day
load("combo_test_v8.Rdata")#File with AURC for each subject on visit 6 on median day
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")#This is the outcome data
##Working with outcomes data 
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()
#By doing it visit by visit I make sure I get those with valid BMI at both visits. 

outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()

##The above are subjects in the sample population with valid BMI at each visit to find those who have BMI values at both visits:
ids_BMI<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1309 have BMI at both visits.

outcomes6<-combo_test_v6%>% #Filter the AURC data to include only those 1309
  dplyr::filter(ID %in%ids_BMI)%>%
  inner_join(outcomes_v6,by="ID")
outcomes8<-combo_test_v8%>%
  dplyr::filter(ID %in%ids_BMI)%>%
  inner_join(outcomes_v8,by="ID")

##Combine it all:
coutcomes6<-chond_outcome%>%
  inner_join(outcomes6,by=c("ID","VISIT"))
coutcome_8<-chond_outcome%>%
  inner_join(outcomes8,by=c("ID","VISIT"))

coutcome<-rbind(coutcomes6,coutcome_8)#Final combined

mixed_model <- mixed_model(
  fixed = chond ~ AURC + AGE + SEX + BMI + RACE,
  random = ~ VISIT||ID,
  family = binomial,
  data = coutcome,control = list(max_coef_value=100, iter_EM = 0))

```

```{r Sup: Osteophyte}
osteo_6<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_sq_bu06.txt", sep="|", header=T)
osteo_8<-read.table("/project/o/oespinga/oespinga/OAI/ClinicalData/CompleteData_ASCII/kxr_sq_bu08.txt", sep="|", header=T)
load("visits_ID.Rdata")#IDs of those in the sample pop 

osteoph_6<-osteo_6%>%
  dplyr::filter(ID %in% visits_ID$ID)%>% #This left us with 1309 subjects I had 1309 unique barcodes
  select(ID,SIDE,V06XROSFL,V06XROSFM,V06XROSTL,V06XROSTM,V06BARCDBU)%>% #I look at only the femur variable
  na.omit()%>% #Remove NA values measured for femur 
  group_by(ID,SIDE)%>%#This leaves us 823 subjects
  dplyr::summarise(V06XROSFL=max(V06XROSFL),
                   V06XROSFM=max(V06XROSFM),
                   V06XROSTL=max(V06XROSTL),
                   V06XROSTM=max(V06XROSTM))%>% #Finds the max measured for each side with repeats
  as.data.frame()%>% #Leaving us with 823 subjects 
  group_by(ID)%>%
  dplyr::summarise(fl=sum(V06XROSFL),
                   fm=sum(V06XROSFM),
                   tl=sum(V06XROSTL),
                    tm=sum(V06XROSTM))%>%
  mutate(osteo=ifelse(fl==0& fm==0&tl==0&tm==0,0,1))%>%
  mutate(VISIT=as.factor(c("V06")))#Making the outcome into a binary where those with 1 on at least one side is graded as a 1 and all else is a 0.

osteoph_8<-osteo_8%>%
  dplyr::filter(ID %in% visits_ID$ID)%>% #This left us with 1309 subjects I had 1309 unique barcodes
  select(ID,SIDE,V08XROSFL,V08XROSFM,V08XROSTL,V08XROSTM,V08BARCDBU)%>% #I look at only the femur variable
  na.omit()%>% #Remove NA values measured for femur 
  group_by(ID,SIDE)%>%#This leaves us 823 subjects
  dplyr::summarise(V08XROSFL=max(V08XROSFL),
                   V08XROSFM=max(V08XROSFM),
                   V08XROSTL=max(V08XROSTL),
                   V08XROSTM=max(V08XROSTM))%>% #Finds the max measured for each side with repeats
  as.data.frame()%>% #Leaving us with 823 subjects 
  group_by(ID)%>%
  dplyr::summarise(fl=sum(V08XROSFL),
                   fm=sum(V08XROSFM),
                   tl=sum(V08XROSTL),
                    tm=sum(V08XROSTM))%>%
  mutate(osteo=ifelse(fl==0& fm==0&tl==0&tm==0,0,1))%>%
  mutate(VISIT=as.factor(c("V08")))#Making the outcome into a binary where those with 1 on at least one side is graded as a 1 and all else is a 0

#I want to check if I have the same IDs at both visits
identical(osteoph_8$ID,osteoph_6$ID) #I get back false

#Get the IDs of people with measurements at both visits 
#Extract list of IDs from each dataset and find shared ID list. 
os_6_ids<-osteoph_6%>%
  select(ID)

os_8_ids<-osteoph_8%>%
  select(ID)

ids_ost<-inner_join(os_6_ids,os_8_ids,by="ID") #Leaving me with 115

##Combine the outcomes for each visit
ost_outcome<-rbind(osteoph_8,osteoph_6)%>%
  dplyr::filter(ID %in% ids_ost$ID)%>%
  group_by(ID)%>%
  as.data.frame()##Final with medial outcomes

load("combo_test_v6.Rdata")#File with AURC for each subject on visit 8 on median day
load("combo_test_v8.Rdata")#File with AURC for each subject on visit 6 on median day
load("/gpfs/fs1/home/o/oespinga/reshani0/OAI_data_long_202202.RData")#This is the outcome data
##Working with outcomes data 
outcomes_v8<-OAI_KL_JSW_data%>% #The outcome data 
  dplyr::filter(VISIT=="V08")%>% #Visit 8 
  dplyr::filter(ID%in% visits_ID$ID)%>% #Get our sample pop IDs only
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()
#By doing it visit by visit I make sure I get those with valid BMI at both visits. 

outcomes_v6<-OAI_KL_JSW_data%>%
  dplyr::filter(VISIT=="V06")%>% #Visit 6
  dplyr::filter(ID%in% visits_ID$ID)%>%
  select(ID,VISIT,AGE,SEX,BMI,RACE)%>%
  transform(VISIT=as.factor(VISIT))%>%
  na.omit()

##The above are subjects in the sample population with valid BMI at each visit to find those who have BMI values at both visits:
ids_BMI<-c(intersect(outcomes_v6$ID,outcomes_v8$ID)) #1309 have BMI at both visits.

outcomes6<-combo_test_v6%>% #Filter the AURC data to include only those 1309
  dplyr::filter(ID %in%ids_BMI)%>%
  inner_join(outcomes_v6,by="ID")
outcomes8<-combo_test_v8%>%
  dplyr::filter(ID %in%ids_BMI)%>%
  inner_join(outcomes_v8,by="ID")

##Combine it all:
ostutcomes6<-ost_outcome%>%
  inner_join(outcomes6,by=c("ID","VISIT"))
ostutcome_8<-ost_outcome%>%
  inner_join(outcomes8,by=c("ID","VISIT"))

ostoutcome<-rbind(ostutcomes6,ostutcome_8)#Final combined


mixed_model <- mixed_model(
  fixed = osteo ~ AURC + AGE + SEX + BMI + RACE,
  random = ~ VISIT|ID,
  family = binomial,
  data = ostoutcome,control = list(max_coef_value=100, iter_EM = 0))

```
